---
title: "yas_ae_linkage_report"
author: "Ric Campbell"
date: "15/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(tinytex)
library(knitr)
library(kableExtra)
library(googlesheets4)
library(googledrive)
library(lubridate)
library(ggplot2)
```

```{r read_in_data, echo=FALSE}

  ##yas_data <- load("D:/SINEPOST/SINEPOSTDM/data/datasets/epr_tables_datetime_changes-2021-07-14-100514.rda")
  ae_data <- readRDS("D:/SINEPOST/SINEPOSTDM/data/datasets/ae_processed_2021-08-02-152502.rds")
  yas_ae_data <- readRDS("D:/SINEPOST/SINEPOSTDM/data/datasets/yas_ae_ecds_linked_2021-09-13-154308.rds")
  study_id_encrypted_hesid_lookup <- readRDS("D:/SINEPOST/SINEPOSTDM/data/linkage/study_id_hesid_lookup_2021-08-02-152440.rds")

```

```{r link_yas_to_ae_data, echo = FALSE}

## Merge epr single table with epr point of care to get hospital information for all eprs (point of care is 1-to-1 to epr)
  
  stopifnot(epr_point_of_care_table[, .N, by = epr_id][N > 1, .N] == 0)
  
  epr_single_value_fields_table_poc <- merge(epr_single_value_fields_table,
                                             epr_point_of_care_table,
                                             by = "epr_id",
                                             all.x = TRUE)
  
  
## Merge in HESID to yas data single table then can used epr_id to link
  
  epr_single_value_fields_table_poc_HES_id <- merge(epr_single_value_fields_table_poc,
                                                    study_id_encrypted_hesid_lookup,
                                                    by.x = "patient_id",
                                                    by.y = "STUDY_ID",
                                                    all.x = TRUE)
  

## Merge in all AE arrival times for each person (HESID)

  yas_hes_linkage <- merge(epr_single_value_fields_table_poc_HES_id[transported_to_hospital == TRUE & receiving_hospital_department %chin% c("ED", "ED - Self Handover", "ED Self-handover/minors","ED Resus"),
                                                                    .(ENCRYPTED_HESID, incident_datetime, epr_id, patient_id)],
                           ae_data[, .(ENCRYPTED_HESID, ARRIVALTIME, AEKEY)],
                           by = "ENCRYPTED_HESID",
                           all.x = TRUE)


## Calculate the time between ambulance incident time, and arrival at A&E

  yas_hes_linkage[, incident_to_arrival_time := as.numeric(difftime(ARRIVALTIME, incident_datetime, unit = "hours"))]
  
  yas_hes_linkage[, incident_to_arrival_time_clipped := pmax(pmin(incident_to_arrival_time, 24.0), -24.0)]
  
  yas_hes_linkage[, incident_to_arrival_time_absolute := abs(incident_to_arrival_time)]


## Set order to find smallest absolute time between ambulance incident time and A&E arrival time - need an extra field to make sure sort is same everytime

  setorder(yas_hes_linkage, incident_to_arrival_time_absolute, na.last = TRUE)
  
  yas_hes_linkage[!is.na(incident_to_arrival_time_absolute), order := 1:.N, by = epr_id]


### Select the smallest absolute time that was within 24 hours, but also remove any cases that were within the first and last day of data

  smallest_time_diff_per_amb_inc_24_hour <- yas_hes_linkage[order == 1 &
                                                              incident_datetime >= floor_date(min(incident_datetime), unit = "day") + days(1) &
                                                              incident_datetime < floor_date(max(incident_datetime), unit = "day")]

  
## Need to order the other way so as to keep AEKEY-epr_id 1-to-1, per_id added to sort to keep same each time
## Will this get rid of epr-(HES, incident_times) multiples, as HES-AEKEY-epr is now all 1-to-1?
  
  setorder(smallest_time_diff_per_amb_inc_24_hour, incident_to_arrival_time_absolute, epr_id, na.last = TRUE)
  
  smallest_time_diff_per_amb_inc_24_hour[!is.na(incident_to_arrival_time_absolute), order := 1:.N, by = AEKEY]
  
  smallest_time_diff_per_amb_inc_24_hour <- smallest_time_diff_per_amb_inc_24_hour[order == 1]  

```

```{r create_numbers_to_report, echo = FALSE}

## AE data on own
## number of distinct patients (HES_ID), and distinct records (AEKEY)
  ae_distinct_records <- uniqueN(ae_data[, AEKEY])
  ae_distinct_patient <- uniqueN(ae_data[, ENCRYPTED_HESID])

# no. records YAS
  num_yas_incidents <- epr_single_value_fields_table[, .N]
  
# no./percent that did/didn't have patient id
  num_yas_incidents_patient_ids <- epr_single_value_fields_table[!is.na(patient_id), .N]
  percent_yas_patient_ids <- round((num_yas_incidents_patient_ids/num_yas_incidents)*100, digits = 2)

# no./percent of STUDY_IDs not traced - YAS
  num_yas_not_traced_study <- uniqueN(epr_single_value_fields_table_poc_HES_id[is.na(ENCRYPTED_HESID) & !is.na(patient_id), patient_id])
  percent_yas_not_traced_study <- round((num_yas_not_traced_study/uniqueN(epr_single_value_fields_table[, patient_id])*100), digits = 2)
  
  
# no./percent of incidents that were traced
  num_yas_traced_patients <- uniqueN(epr_single_value_fields_table_poc_HES_id[!is.na(ENCRYPTED_HESID), ENCRYPTED_HESID])
  num_yas_traced_incidents <- uniqueN(epr_single_value_fields_table_poc_HES_id[!is.na(ENCRYPTED_HESID), epr_id])
  percent_yas_traced_incidents <- round((uniqueN(epr_single_value_fields_table_poc_HES_id[!is.na(ENCRYPTED_HESID), epr_id])*100)/num_yas_incidents, digits = 2)
  percent_yas_patient_ids_traced_incidents <- round((uniqueN(epr_single_value_fields_table_poc_HES_id[!is.na(ENCRYPTED_HESID) & !is.na(patient_id),
                                                                                               epr_id])*100)/num_yas_incidents_patient_ids, digits = 2)
  

# YAS incidents conveyed to hospital
  
  yas_conveyed <- epr_single_value_fields_table_poc_HES_id[!is.na(ENCRYPTED_HESID) & transported_to_hospital == TRUE & receiving_hospital_department %chin% c("ED", "ED - Self Handover", "ED Self-handover/minors","ED Resus")]

  num_yas_incidents_patient_ids_conveyed <- yas_conveyed[!is.na(patient_id), .N]
  percent_yas_patient_id_incidents_converyed <- round((yas_conveyed[!is.na(patient_id),
                                                                    .N]/epr_single_value_fields_table_poc_HES_id[!is.na(ENCRYPTED_HESID) &!is.na(patient_id), .N])*100, digits = 2)
  

## YAS-AE linked

# Incidents that have AE data
  
  number_incidents_ae_data <- uniqueN(yas_ae_data[!is.na(AEKEY), epr_id])
  percentage_incidents_ae_data <- round((uniqueN(yas_ae_data[!is.na(AEKEY), epr_id])*100)/num_yas_incidents_patient_ids_conveyed, digits = 2)
  
  
## Incidents that were in the first data of YAS data that were linked

  incidents_removed_first_day <- yas_hes_linkage[order == 1 & 
                                              incident_datetime < floor_date(min(incident_datetime), unit = "day") + days(1), .N]
  
  incidents_removed_last_day <- yas_hes_linkage[order == 1 & 
                                              incident_datetime >= floor_date(max(incident_datetime), unit = "day"), .N]
  
  
  
### Distribution of epr-AE linked records within 24 of each other

  # ggplot(smallest_time_diff_per_amb_inc_24_hour[incident_to_arrival_time_absolute <= 5], aes(x = incident_to_arrival_time_clipped)) +
  #   geom_histogram(binwidth = 1) +
  #   ggtitle("Distribution of time between ambulance incident and A&E arrival")


```

```{r put_tables_together, echo = FALSE}
## To report on:

## Reattenders
## Low Acuity numbers - use function



# how many record in first day of YAS data that were useful
# how many in last 6 hours of YAS data that were useful

```

## AE Data


Table that shows the number of records and patients (unique HES_IDs) present in the AE data received from NHS Digital

```{r ae_tables, echo=FALSE}

ae_table <- data.table(number_of_records = ae_distinct_records, number_of_patients = ae_distinct_patient)

kable_styling(kable(ae_table, format = "html"))

```



## YAS Data


Table that shows the number/percentage of incidents (epr_ids) present in the ambulance data received from YAS.  
Patient ids were created in the YAS data if the record contained either a NHS number or all three of date of birth, sex and incident postcode.  
(N.B - Patient ids in the YAS data and do not always relate to a single unique patient, they are are named Study ids in the NHSD data.

```{r yas_tables1, echo=FALSE}

yas_table_1 <- data.table(number_yas_incidents = num_yas_incidents, number_of_incidents_with_patient_id = num_yas_incidents_patient_ids,
                        percentage_of_incidents_with_patient_id = percent_yas_patient_ids)
  
kable_styling(kable(yas_table_1, format = "html"))


```


Table showing the number/percentage of patient ids in the YAS data that were not traced by NHS Digital.  
A patient id was considered to be traced if a corresponding HES id was returned in the AE dataset.  

```{r yas_tables2, echo=FALSE}

yas_table_2 <- data.table(number_yas_patient_ids_not_traced = num_yas_not_traced_study,
                          percentage_yas_patient_ids_not_traced = percent_yas_not_traced_study)
  
kable_styling(kable(yas_table_2, format = "html"))


```



## YAS data that was traced


Table showing the number of patients and incidents that have been traced in the YAS data.  
A patient id was considered to be traced if a corresponding HES id was returned in the AE dataset.  
  
Percentages show the number of incidents traced related to:  
  1) the total number of incidents in the YAS dataset (including those without patient ids)  
  2) the number of incidents where a patient id was present, therefore representing the number of incidents it was theoretically possible to trace  


```{r yas_tables3, echo=FALSE}

yas_table_3 <- data.table(number_of_patients_traced_in_yas_data = num_yas_traced_patients,
                          number_of_incidents_traced = num_yas_traced_incidents,
                          percentage_of_all_incidents_traced = percent_yas_traced_incidents,
                          percentage_of_incidents_with_patient_id_traced = percent_yas_patient_ids_traced_incidents)
  
kable_styling(kable(yas_table_3, format = "html"))


```



### Conveyance


Table showing the number of incidents **that were traced** in the YAS dataset, that stated the patient was transported to the Emergency Department.  
A patient was considered to have been conveyed to AE if:  
  transported_to_hospital(poctransported) == TRUE  
    AND   
  receiving_hospital_department (pocdepartment) referred to an ED  
  
Percentages show the number of incidents traced related to the number of incidents where a patient id was present, therefore representing the number of incidents it was theoretically possible to trace  


```{r yas_tables4, echo=FALSE}

yas_table_4 <- data.table(number_of_incidents_with_patient_id_conveyed = num_yas_incidents_patient_ids_conveyed,
                          percentage_of_incidents_with_patients_id_stated_were_conveyed = percent_yas_patient_id_incidents_converyed)
  
kable_styling(kable(yas_table_4, format = "html"))


```



### YAS-AE data (within 0 - 6 hours of ambulance arrival time)


Table showing the number and percentage of ambulance incidents **that stated the patient was conveyed** that have a AE record attached to them.  
All ambulance incidents (where possible) were linked to only the most recent AE record using absolute time.  
All AE records were linked to only the most recent ambulance incident.  
This means that single ambulance incident has been linked to a single AE record.  
Percentage is in relation to the number of incidents that were traced.  

```{r yas_tables5, echo=FALSE}

yas_table_5 <- data.table(number_of_incidents_that_have_AE_data = number_incidents_ae_data,
                          percentage_of_traced_incidents_that_have_AE_data = percentage_incidents_ae_data)
                          
kable_styling(kable(yas_table_5, format = "html"))


```



## Incidents that were in the first and last day of YAS data that were linked


Incidents that were on the first and last day of the YAS data were removed due to YAS and NHSD data having the same start and end date.    
A full 24 hours either side of the incident datetime was required to check for the closest AE record.  
(Numbers below have already been removed from the counts in the table above).  

```{r yas_table6, echo = FALSE}

yas_table_6 <- data.table(number_of_incidents_removed_on_first_day_of_data = incidents_removed_first_day,
                          number_of_incidents_removed_on_last_day_of_data = incidents_removed_last_day)
  
kable_styling(kable(yas_table_6, format = "html"))

```
