---
title: "NHS Digital Data Checking"
author: "Ric Campbell"
date: "2021-06-29"
output: html_document
---

## Internal UoS PRIEST team only, please do not share more widely

```{r get google login access, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(tinytex)
library(knitr)
library(kableExtra)
library(googlesheets4)
library(googledrive)
library(lubridate)
library(ggplot2)


fn_covertStrToDate <- function(date_in, dt_format = "%Y-%m-%d") {
  date_out <- as.Date(lubridate::fast_strptime(paste(date_in, "12:00:00"), format = paste(dt_format, "%H:%M:%S"), tz = "Europe/London", lt = FALSE))
  date_out <- replace(date_out, date_out %in% as.Date("1900-01-01"), NA)
  return(date_out)
}

fn_removeBlanks <- function(field) {
  field <- trimws(field)
  return(replace(field, field == "", NA))
}


## Give script permission to use auth token found in cache

warning("You must have connected to your Google account proir to running this script as it uses the auth token found in the cache")
# gs4_auth() - may need re-doing even if have not shut down RStudio
# Then run markdown from console:
# rmarkdown::render("RMarkdown/nhs_digital_data_checking_and_outputs.Rmd")

options(gargle_oauth_email = TRUE)

```


```{r, read in data, echo = FALSE}

## Read in all data from NHS Digital (and YAS data for percentages later)

ecds <- fread("D:/data/nhs_digital/ecds/FILE0119464_NIC284866_HES_ECDS_201999.txt",
              colClasses = "character",
              na.strings = "")

ae <- fread("D:/data/nhs_digital/ae/FILE0119465_NIC284866_HES_AE_201999.txt",
            colClasses = "character",
            na.strings = "")

yas <- data.table(readRDS("D:/data/yas/sinepost.rds"))


## Read in receipt file for extra numbers

ecds_receipt <- fread("D:/data/nhs_digital/ecds/FILE0119464_NIC284866_HES_ECDS_201999.txt_CPOS-631961_Completed.csv",
                      colClasses = "character")
ae_receipt <- fread("D:/data/nhs_digital/ae/FILE0119465_NIC284866_HES_AE_201999.txt_CPOS-626812_Completed.csv",
                    colClasses = "character")

```


```{r, create number for table, echo = FALSE}

# Convert date fields to dates

ecds[, ARRIVAL_DATE := fn_covertStrToDate(ARRIVAL_DATE)]
ae[, ARRIVALDATE := fn_covertStrToDate(ARRIVALDATE)]
yas[, incdate := as.Date(incDate, format = "%Y-%m-%d")]


## Read in col names that we asked for from Google Sheet

nhs_digital_requested_data <- as.character(as_sheets_id("https://https://docs.google.com/spreadsheets/d/1djY-ldYUVBwUUcmso1quWkn3NebpIm3Hrx74WASlVRs/edit#gid=583528663"))


## Read in sites sheet from google drive - correctly formatted field name has been done

## Might have to put a delay in as too many read requests in short space of time?

suppressMessages(ecds_requested <- data.table(range_read(nhs_digital_requested_data,
                                                         sheet = "ECDS", col_types = "c")))

suppressMessages(ae_requested <- data.table(range_read(nhs_digital_requested_data,
                                                       sheet = "AE", col_types = "c")))


## Change all col names to upper case and take col names

ecds_colnames <- colnames(setnames(ecds, colnames(ecds), toupper(colnames(ecds))))
ae_colnames <- colnames(setnames(ae, colnames(ae), toupper(colnames(ae))))


## Make changes to requested list to add entries that have multiple fields

## ECDS - number of fields that should be present created to match up with data provided
## All fields have a group (or field) repeat of 1...* as found in the technical output specification (V3.0) available on NHS Digital website

ecds_requested_names <- c(ecds_requested[field_name != "COMORBIDITIES" & field_name != "COMORBIDITIES_VALID_APPROVED" & field_name != "DIAGNOSIS_CODE"
                                         & field_name != "DIAGNOSIS_FIRST" & field_name != "DIAGNOSIS_IS_AEC_RELATED" & field_name != "DIAGNOSIS_IS_ALLERGY_RELATED"
                                         & field_name != "DIAGNOSIS_IS_APPLICABLE_TO_FEMALES" & field_name != "DIAGNOSIS_IS_APPLICABLE_TO_MALES"
                                         & field_name != "DIAGNOSIS_IS_INJURY_RELATED" & field_name != "DIAGNOSIS_IS_NOTIFIABLE_DISEASE"
                                         & field_name != "DIAGNOSIS_QUALIFIER" & field_name != "DIAGNOSIS_QUALIFIER_IS_APPROVED" 
                                         & field_name != "DIAGNOSIS_VALID_APPROVED" & field_name != "DRUG_ALCOHOL_CODE" 
                                         & field_name != "DRUG_ALCOHOL_CODE_VALID_APPROVED" & field_name != "INVESTIGATION_CODE" 
                                         & field_name != "INVESTIGATION_DATE" & field_name != "INVESTIGATION_TIME" & field_name != "INVESTIGATIONS_VALID_APPROVED"
                                         & field_name != "REFERRAL_ASSESSMENT_DATE" & field_name != "REFERRED_TO_SERVICE" 
                                         & field_name != "REFERRED_TO_SERVICE_VALID_APPROVED" & field_name != "SEQUENCE_NUMBER"
                                         & field_name != "SERVICE_REQUEST_DATE" & field_name != "SERVICE_REQUEST_TIME" & field_name != "TREATMENT_CODE"
                                         & field_name != "TREATMENT_DATE" & field_name != "TREATMENT_TIME" & field_name != "TREATMENTS_VALID_APPROVED", field_name],
                          paste0("COMORBIDITIES_", 1:10), paste0("COMORBIDITIES_VALID_APPROVED_", 1:10), paste0("DIAGNOSIS_CODE_", 1:12),
                          paste0("DIAGNOSIS_FIRST_", 1:12), paste0("DIAGNOSIS_IS_AEC_RELATED_", 1:12),paste0("DIAGNOSIS_IS_ALLERGY_RELATED_", 1:12),
                          paste0("DIAGNOSIS_IS_APPLICABLE_TO_FEMALES_", 1:12), paste0("DIAGNOSIS_IS_APPLICABLE_TO_MALES_", 1:12),
                          paste0("DIAGNOSIS_IS_INJURY_RELATED_", 1:12), paste0("DIAGNOSIS_IS_NOTIFIABLE_DISEASE_", 1:12), paste0("DIAGNOSIS_QUALIFIER_", 1:12),
                          paste0("DIAGNOSIS_QUALIFIER_IS_APPROVED_", 1:12), paste0("DIAGNOSIS_VALID_APPROVED_", 1:12), paste0("DRUG_ALCOHOL_CODE_", 1:4),
                          paste0("DRUG_ALCOHOL_CODE_VALID_APPROVED_", 1:4), paste0("INVESTIGATION_CODE_", 1:12), paste0("INVESTIGATION_DATE_", 1:12),
                          paste0("INVESTIGATION_TIME_", 1:12), paste0("INVESTIGATIONS_VALID_APPROVED_", 1:12), paste0("REFERRAL_ASSESSMENT_DATE_", 1:4),
                          paste0("REFERRED_TO_SERVICE_", 1:4), paste0("REFERRED_TO_SERVICE_VALID_APPROVED_", 1:4), paste0("SEQUENCE_NUMBER_", 1:12),
                          paste0("SERVICE_REQUEST_DATE_", 1:4), paste0("SERVICE_REQUEST_TIME_", 1:4), paste0("TREATMENT_CODE_", 1:12),
                          paste0("TREATMENT_DATE_", 1:12), paste0("TREATMENT_TIME_", 1:12), paste0("TREATMENTS_VALID_APPROVED_", 1:12))


## AE - number of fields that should be present created to match up with data provided
## Number of fields that should be present taken from data dictionary ("only the first 12 [X] are available within HES")

ae_requested_names <- c(ae_requested[field_name != "DIAG_NN" & field_name != "DIAG2_NN" & field_name != "DIAG3_NN" & field_name != "DIAGA_NN" 
                                     & field_name != "DIAGS_NN" & field_name != "INVEST_NN" & field_name != "INVEST2_NN" & field_name != "TREAT_NN" 
                                     & field_name != "TREAT2_NN" & field_name != "TREAT3_NN", field_name],
                        paste0("DIAG_", formatC(1:12, width = 2, format = "d", flag = "0")), paste0("DIAG2_", formatC(1:12, width = 2, format = "d", flag = "0")),
                        paste0("DIAG3_", formatC(1:12, width = 2, format = "d", flag = "0")), paste0("DIAGA_", formatC(1:12, width = 2, format = "d", flag = "0")),
                        paste0("DIAGS_", formatC(1:12, width = 2, format = "d", flag = "0")), paste0("INVEST_", formatC(1:12, width = 2, format = "d", flag = "0")),
                        paste0("INVEST2_", formatC(1:12, width = 2, format = "d", flag = "0")), paste0("TREAT_", formatC(1:12, width = 2, format = "d", flag = "0")),
                        paste0("TREAT2_", formatC(1:12, width = 2, format = "d", flag = "0")), paste0("TREAT3_", formatC(1:12, width = 2, format = "d", flag = "0")))


## Check that we have received all the cols asked for, and if have received any more

missing_ecds_cols <- ecds_requested_names[!ecds_requested_names %in% ecds_colnames]
extra_ecds_cols <- ecds_colnames[!ecds_colnames %in% ecds_requested_names]

missing_ae_cols <- ae_requested_names[!ae_requested_names %in% ae_colnames]
extra_ae_cols <- ae_colnames[!ae_colnames %in% ae_requested_names]


## Get list on unique patient_ids from YAS data to see how many are in each data set

ecds_patients <- unique(ecds[, STUDY_ID])
ae_patients <- unique(ae[, STUDY_ID])
yas_patients <- unique(yas[, patient_id])


## Find number of unique records, unique study_ids (or similar) for each dataset before building table

### EDCS

ecds_records <- unique(ecds)[, .N]
ecds_study_ids <- length(ecds[!is.na(STUDY_ID), unique(STUDY_ID)])
yas_study_ids_in_ecds_percent <- round(uniqueN(yas[patient_id %in% ecds_patients, patient_id])/length(yas_patients)*100, digits = 2)

ecds_study_ids_in_yas_percent <- round(uniqueN(ecds[STUDY_ID %in% yas_patients, STUDY_ID])/ecds_study_ids*100, digits = 2)


### AE
## number of records when KEY fields are removed?
## all_non_key_cols <- ae_requested_names[!ae_requested_names %in% "AEKEY"]

ae_records <- unique(ae)[, .N]
ae_study_ids <- length(ae[!is.na(STUDY_ID), unique(STUDY_ID)])
ae_study_ids_percent <- round(uniqueN(yas[patient_id %in% ae_patients, patient_id])/length(yas_patients)*100, digits = 2)

ae_study_ids_in_yas_percent <- round(uniqueN(ae[STUDY_ID %in% yas_patients, STUDY_ID])/ae_study_ids*100, digits = 2)


### YAS

yas_records <- unique(yas)[, .N]
yas_study_ids <- length(yas[!is.na(patient_id), unique(patient_id)])

yas_records_not_traced <- yas[(!patient_id %in% ecds_patients) & !(patient_id %in% ae_patients)]
yas_patients_not_traced <- uniqueN(yas_records_not_traced[!is.na(patient_id), patient_id])


## Percent of study ids that are in both datasets?


## Create table with numbers of rows, distinct records, unique study_ids, min/max dates

dataset_numbers <- data.table(dataset = "ECDS", number_of_rows = ecds[, .N], number_of_records = ecds_records, 
                              number_study_ids = ecds_study_ids, percent_of_yas_study_ids_present = yas_study_ids_in_ecds_percent,
                              percent_study_ids_in_yas = ae_study_ids_in_yas_percent, number_of_rows_removed_nhsd = as.integer(ecds_receipt$`Number of Rows Removed`),
                              number_of_unique_patients_removed_nhsd = as.integer(ecds_receipt$`Number of Unique Patients Removed`),
                              min_date = as.character(ecds[!is.na(ARRIVAL_DATE), min(ARRIVAL_DATE)]),
                              max_date = as.character(ecds[!is.na(ARRIVAL_DATE), max(ARRIVAL_DATE)]))

dataset_numbers <- rbind(dataset_numbers, list("A&E", ae[, .N], ae_records, ae_study_ids, ae_study_ids_percent, ae_study_ids_in_yas_percent,
                                               as.integer(ae_receipt$`Number of Rows Removed`), as.integer(ae_receipt$`Number of Unique Patients Removed`), 
                                               as.character(ae[!is.na(ARRIVALDATE), min(ARRIVALDATE)]),
                                               as.character(ae[!is.na(ARRIVALDATE), max(ARRIVALDATE)])))

dataset_numbers <- rbind(dataset_numbers, list("YAS", yas[, .N], yas_records, yas_study_ids, 100L, 100L, NA, NA,
                                               as.character(yas[!is.na(incdate), min(incdate)]),
                                               as.character(yas[!is.na(incdate), max(incdate)])))

```


``` {r duplicate check and graph creation, echo = FALSE}

## Separate full duplicate

ecds[, duplicated_record := duplicated(ecds)]
ae[, duplicated_record := duplicated(ae)]

ecds_data_duplicates <- ecds[duplicated_record == TRUE]
ecds_no_duplicates <- ecds[duplicated_record == FALSE]
ae_data_duplicates <- ae[duplicated_record == TRUE]
ae_no_duplicates <- ae[duplicated_record == FALSE]


ecds_number_of_full_dups <- ecds[, .N] - ecds_no_duplicates[ ,.N]
ecds_percentage_full_dups <- (ecds_number_of_full_dups/ecds[, .N])*100L
ae_number_of_full_dups <- ae[, .N] - ae_no_duplicates[ ,.N]
ae_percentage_full_dups <- (ae_number_of_full_dups/ae[, .N])*100L


## Change col types as needed

ecds_no_duplicates[, AGE_AT_ARRIVAL := as.integer(AGE_AT_ARRIVAL)]
ae_no_duplicates[, ACTIVAGE := as.integer(ACTIVAGE)]


## Floor dates for plotting

ecds_no_duplicates[, incident_month := floor_date(ARRIVAL_DATE, unit = "month")]
ecds_no_duplicates[, incident_week := floor_date(ARRIVAL_DATE, unit = "week")]
ae_no_duplicates[, incident_month := floor_date(ARRIVALDATE, unit = "month")]
ae_no_duplicates[, incident_week := floor_date(ARRIVALDATE, unit = "week")]


## Create age groups

ecds_no_duplicates[, age_group := cut(AGE_AT_ARRIVAL, c(seq(0, 120, 10), Inf), include.lowest = TRUE)]  
ae_no_duplicates[, age_group := cut(ACTIVAGE, c(seq(0, 120, 10), Inf), include.lowest = TRUE)]


## Create a few graphs of records over time

## ECDS

ecds_records_per_month_plot <- ggplot(ecds_no_duplicates[, .N, by = incident_month],
                                      aes(x = incident_month, y = N)) +
  geom_point() + 
  geom_smooth() + 
  ggtitle("ECDS episodes per month")

ecds_records_per_week_plot <- ggplot(ecds_no_duplicates[, .N, by = incident_week],
                                     aes(x = incident_week, y = N)) +
  geom_point() + 
  geom_smooth() + 
  ggtitle("ECDS episodes per week")

ecds_age_plot <- ggplot(ecds_no_duplicates[, .N, by = age_group],
                        aes(x = age_group, y = N)) +
  geom_point() + 
  geom_smooth() + 
  ggtitle("ECDS episodes per age group")

ecds_age_month_plot <- ggplot(ecds_no_duplicates[, .N, by = .(incident_month, age_group)],
                              aes(x = incident_month, y = N, color = age_group)) +
  geom_point() + 
  geom_smooth() + 
  ggtitle("ECDS episodes by age group per month")


## AE

ae_records_per_month_plot <- ggplot(ae_no_duplicates[, .N, by = incident_month],
                                    aes(x = incident_month, y = N)) +
  geom_point() + 
  geom_smooth() + 
  ggtitle("AE episodes per month")

ae_records_per_week_plot <- ggplot(ae_no_duplicates[, .N, by = incident_week],
                                   aes(x = incident_week, y = N)) +
  geom_point() + 
  geom_smooth() + 
  ggtitle("AE episodes per week")

ae_age_plot <- ggplot(ae_no_duplicates[, .N, by = age_group],
                      aes(x = age_group, y = N)) +
  geom_point() + 
  geom_smooth() + 
  ggtitle("AE episodes per age group")

ae_age_month_plot <- ggplot(ae_no_duplicates[, .N, by = .(incident_month, age_group)],
                            aes(x = incident_month, y = N, color = age_group)) +
  geom_point() + 
  geom_smooth() + 
  ggtitle("AE episodes by age group per month")


```

# Comparison of the columns that we asked for and we have received for NHS Digital datasets

Missing ECDS columns: `r paste0(length(missing_ecds_cols), collapse = " ")`
Extra ECDS columns: `r paste0(extra_ecds_cols, collapse = " ")`

Missing A&E columns: `r paste0(length(missing_ae_cols), collapse = " ")`  
Extra A&E columns: `r paste0(extra_ae_cols, collapse = " ")`

# Summary of the data received from NHS Digital

```{r dataset numbers, echo = FALSE}
kable_styling(kable(dataset_numbers, format = "html"))

```

Number of YAS records that do not have a patient ID assigned: `r paste0(yas[is.na(patient_id), .N], collapse = " ")`  
Number of YAS records that have not been traced by NHS Digital (patient not found or patient ID not assigned): `r paste0(yas_records_not_traced[, .N], collapse = " ")`  
Number of YAS patients that have not been traced by NHS Digital: `r paste0(yas_patients_not_traced, collapse = " ")`


# Graphs of ECDS data plotted against arrival date
```{r ecds graphs, echo = FALSE}

ecds_records_per_month_plot

ecds_records_per_week_plot

ecds_age_plot

ecds_age_month_plot

```

# Graphs of AE data plotted against arrival date
``` {r ae graphs, echo = FALSE}

ae_records_per_month_plot

ae_records_per_week_plot

ae_age_plot

ae_age_month_plot

```


